---
title: "CPBS7602 Multiple Testing & Permutations"
author: "Lauren Vanderlinden"
date: "Report generated: `r paste(format(Sys.Date(), '%B %d, %Y'))`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, include=FALSE}
rm(list=ls())

library(ggplot2)
library(dplyr)
library(tidyverse)
library(kableExtra)
```

## Multiple Testing 

### Null Hypothesis 

If the null hypothesis is true and there are no true differences in our populations, what would the distributions of p-values look like?

```{r}

# simulated data for a 2-sample t-test
ngenes <- 5000
nsamples <- 5

pvals.null <- apply(as.matrix(1:ngenes), 
               1, function(a) 
                 t.test(rnorm(n = nsamples),
                   rnorm(n = nsamples), 
                    var.equal = TRUE)$p.value)


ggplot(as.data.frame(pvals.null), aes(x = pvals.null)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "darkred") +
  labs(title = "Histogram of p-values No True Effect",
       x = "p-values") +
  theme_minimal(base_size=16)
```


### Alternative Hypothesis 

If the alternative hypothesis is true and there are true differences in our populations, what would the distributions of p-values look like?


```{r}
# simulated data for a 2-sample t-test
ngenes <- 5000
nsamples <- 5
effectsize <- 1

pvals <- apply(as.matrix(1:ngenes), 
               1, function(a) 
                 t.test(rnorm(n = nsamples),
                   rnorm(n = nsamples, mean=effectsize), 
                    var.equal = TRUE)$p.value)

ggplot(as.data.frame(pvals), aes(x = pvals)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "steelblue") +
  labs(title = "Histogram of p-values True Effect Size of 2",
       x = "p-values") +
  theme_minimal(base_size=16)
```

## Q-Q Plots

```{r}
# get a mix of p-values where 15% come from true effect and 85% null  
pvals.mix <- c(sample(x=pvals, size=0.05*ngenes), 
               sample(x=pvals.null, size=0.95*ngenes))

# Order you p-values 
pvals.mix <- pvals.mix[order(pvals.mix)]

# Calculate theoretical quantiles
n <- length(pvals.mix)
theoretical_quantiles <- -log10((1:n) / n)

# Create QQ plot
ggplot(data = NULL, aes(x = theoretical_quantiles, y = -log10(pvals.mix))) +
  geom_point(color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Theoretical Quantiles",
       y = "Observed -log10(P-values)",
       title = "QQ Plot: Mix of True/Null Effect") +
  theme_minimal()+
   theme(text=element_text(size=20))+
  geom_vline(xintercept = median(theoretical.quantiles), color = "red", size=0.5)


median(theoretical_quantiles)
median(-log10(pvals.mix))
```

### Genomic Inflation Factor 


```{r}
# Function to calculate lambda
getLambda = function(pvals){
  chisq <- qchisq(1-pvals,1)
  #Calculate lambda gc (λgc)
  lambda = median(chisq)/qchisq(0.5,1)
  return(lambda)
}

getLambda(pvals.mix)
```